Mergeable Types

### 2017-02-28 paper/examples.tex
edits

The \drawsome example demonstrates the utility of the \name
programming model in building concurrent applications with conceptual
state sharing that make them suitable for transparent deployment in a
distributed setting. The model allows concurrent computations to be
composed around any mergeable type. We now present a series of
examples that demonstrate how mergeable types are built: either by
defining merge functions for ML data types, or by composing existing
mergeable types.

@@

A rope is a data structure that facilitates efficient implementations
of operations such as concatenation, splitting, lookup and insertion
at a random position etc., for a list of values in a distributed
setting. The values are usually characters, with their lists being
strings. The data structure is a binary tree, whose leaf nodes store
substrings of a larger string such that their left-to-right (in-order)
concatenation yeilds the larger string. The idea is to facilitate
efficient lookups and insertions at random positions in the string by
storing position indexes in the internal nodes of the tree.

@@

A non-trivial demonstration illustrating the compositional power of
mergeable types is a shopping list application that allows its users
to collaboratively build a shopping list (listing items in the
decreasing order of their priority), by adding new items with a
specified quantity, deleting existing items, and increasing or
decreasing their quantity. An item (\C{Item.t}) is represented as a
tuple of its name (a string) and the quantity (A \C{Qty.t}).  The
merge function for \C{Item.t} merges the quantities of items (using
\C{Qty.merge}) having the same names. Items with different names are
considered distinct. A shopping list is a mergeable list of items,
i.e., \C{MList(Item)} (a grocery shopping list may not be large enough
to justify using ropes, but \C{MRope(MList(Item))} is certainly
allowed).  \C{MList(Item).merge} automatically lifts the merge
semantics for items to a merge semantics for lists of items.
\begin{wrapfigure}{r}{.35\textwidth}
  \centering
  \includegraphics[width=0.25\textwidth]{Figures/shopping-list}
  \caption{Merging concurrent shopping lists}
  \label{fig:shopping-list}
  \vspace*{-.35in}
\end{wrapfigure}
Fig.~\ref{fig:shopping-list} illustrates its behavior.  Here, two
users, Alice and Bob, concurrently edit a shopping list whose initial
contents are \C{milk} and \C{eggs}. While both simultaneously update
the quantity of eggs, Bob also removes \C{milk} and inserts \C{candy}.
Fig.~\ref{fig:shopping-list-code} shows the \name code for two
concurrent sessions. The merged shopping list contains \C{eggs} and
\C{candy}, where the quantity of eggs is obtained by merging Alice's
and Bob's quantity as per the definition of \C{Qty.merge}.
The examples described above make certain assumptions about the
concurrency model and the underlying system, such as the existence of
a single lowest common ancestor (LCA) for any pair of versions, the
ability to access a previous version on any branch, and mergeability
of any two concurrent versions. Enforcing these guarantees in a fully
decentralized distributed setting requires addressing non-trivial
theoretical challenges. These, and other details that contribute to
the practicality of our model, such as containing the complexity of
merges, are discussed in subsequent sections.

>>>>

The \drawsome example demonstrates the \name's utility in building
concurrent applications with conceptual state sharing that make them
suitable for transparent deployment in a distributed setting. The
model allows concurrent computations to be composed around any
mergeable type. We now present a series of examples that demonstrate
how mergeable types are built - either by defining merge functions for
ML data types, or by composing existing mergeable types.

@@

A rope is a data structure that facilitates efficient implementations
of operations such as concatenation, splitting, lookup and insertion
at a random position etc., for a list of values in a distributed
setting. The values are usually characters, with their lists being
strings. The data structure is a binary tree, whose leaf nodes store
substrings of a larger string such that their left-to-right (in-order)
concatenation yields the larger string. The idea is to facilitate
efficient lookups and insertions at random positions in the string by
storing position indexes in the internal nodes of the tree.

@@

A non-trivial demonstration illustrating the compositional power of
mergeable types is a shopping list application that allows its users
to collaboratively build a shopping list (listing items in the
decreasing order of their priority), by adding new items with a
specified quantity, deleting existing items, and increasing or
decreasing their quantity. An item (\C{Item.t}) is represented as a
tuple of its name (a string) and the quantity (A \C{Qty.t}).  The
merge function for \C{Item.t} merges the quantities of items (using
\C{Qty.merge}) having the same names. Items with different names are
considered distinct. A shopping list is a mergeable list of items,
i.e., \C{MList(Item)} (although a grocery shopping list may not be
large enough to justify using ropes, \C{MRope(MList(Item))} is
certainly allowed).  \C{MList(Item).merge} automatically lifts the
merge semantics for items to a merge semantics for lists of items.
Fig.~\ref{fig:shopping-list} illustrates its behavior.  Here, two
users, Alice and Bob, concurrently edit a shopping list whose initial
contents are \C{milk} and \C{eggs}. While both simultaneously update
the quantity of eggs, Bob also removes \C{milk} and inserts \C{candy}.
Fig.~\ref{fig:shopping-list-code} shows the \name code for two
concurrent sessions. The merged shopping list contains \C{eggs} and
\C{candy}, where the quantity of eggs is obtained by merging Alice's
\begin{wrapfigure}{r}{.35\textwidth}
  \centering
  \includegraphics[width=0.25\textwidth]{Figures/shopping-list}
  \caption{Merging concurrent shopping lists}
  \label{fig:shopping-list}
  \vspace*{.25in}
\end{wrapfigure}
and Bob's quantity as per the definition of \C{Qty.merge}.
The examples described above make certain assumptions about the
concurrency model and the underlying system, such as the existence of
a single lowest common ancestor (LCA) for any pair of versions, the
ability to access a previous version on any branch, and mergeability
of any two concurrent versions. Enforcing these guarantees in a fully
decentralized distributed setting requires addressing non-trivial
theoretical challenges. These, and other details that contribute to
the practicality of our model, such as containing the complexity of
merges, are discussed in subsequent sections.

### 2017-02-28 paper/implementation.tex
edits

We describe the realization of \name programming model on top of a persistent
distributed store and the language extensions to enable and enforce the
constraints of the \name.

>>>>

We describe the realization of \name on top of a persistent
distributed store and the language extensions to enable and enforce
the constraints of the \name.

### 2017-02-28 paper/intro.tex
edits

To illustrate how the kinds of problems described above manifest in
practice, consider how we might write a simple counter library (see
Fig.~\ref{fig:counter-adt}).  A \C{Counter} supports two (update)
operations - \C{add} and \C{mult} - that lets a non-negative integer
value be added or multiplied to the counter, resp.  Observe that the
library is written in an idiomatic functional style, with no special
reasoning principles needed to realize desired functionality.  As long
as applications use the library on a single machine, this
implementation behaves as expected.  However, if the library is used
in the context of a more sophisticated application, say one whose
computation is distributed among a collection of machines, its
behavior can become significantly harder to understand.  In
particular, a distributed implementation might wish to
\emph{replicate} the counter state on each node to improve response
time or fault tolerance.  Unfortunately, adding replication doesn't
come for free.  Attempting to update every replicated copy atomically
is problematic in the absence of sophisticated transaction support,
which impose significant performance penalties.  But, without such
heavyweight mechanisms, applying an \C{Add} operation on one node may
not be instantaneously witnessed on another, which may be in the
process of simultaneously attempting to perform its own \C{Add} or
\C{Mult} action.  While synchronizing the activities of all nodes to
ensure at most one such operation is performed at a time is
impractical, designating a single node to hold the counter state
eschewing replication altogether (as in a client-server
configuration~\cite{Armstrong}), is also an undesirable solution,
given the sensitivity of such architectures to network partitions and
server failures, and the negative performance impact it incurs in
geo-distributed environments~\cite{Walter}.  Removing coordination
altogether by simply replicating counter state without having any
supporting consistency protocol is an equally infeasible approach.

@@

	\vspace*{-1.2in}

@@

Our main innovation is thus the development of a programming model
that realizes a monadic version control system centered around data,
rather than file, coherence.  The \name (Versioned ML) monad lets
programmers write and compose concurrent computations around multiple
(implicit) versions of a mergeable datatype, an ordinary ML datatype
additionally equipped with a \emph{merge} function responsible for
deriving a consistent global state from a collection of local versions
of that state.  A computation progresses by forking (i.e.,
replicating) new versions of existing versions, allowing
synchronization-free local computation to proceed on those nodes,
creating new versions along an existing \emph{branch} that represents
updated local instances, and by merging branches to realize global
consistency.

Our model allows ML programmers to get the benefits of achieving
highly available (low-latency) distributed computation, while
continuing to enjoy the comforts of high-level reasoning and the
familiarity of using standard libraries already provided by the
language.  Notably, while \name's programming model makes no explicit
reference to any specific operational manifestation of a distributed
system (e.g., programmers do not need to explicitly manage replicas),
we demonstrate that it can be nonetheless efficiently realized on
existing real-world geo-replicated distributed systems.

@@

    \item We describe \name programming model that brings to bear the
      power and flexibility of a version control system to the
      administration of replicated data. \name hides the complexity of
      version control behind a monad, and exposes its functionality
      via a simple API that lets programmers define and compose
      distributed ML computations around such data.  Issues related to
      replication only manifest implicitly in the definition of a
      merge operation that defines coherence among different instances
      of replicated state.

>>>>

To illustrate how the kinds of problems described above manifest in
practice, consider how we might write a simple counter library (see
Fig.~\ref{fig:counter-adt}).  A \C{Counter} supports two (update)
operations - \C{add} and \C{mult} - that lets a non-negative integer
value be added or multiplied to the counter, resp.  Observe that the
library is written in an idiomatic functional style, with no special
reasoning principles needed to realize desired functionality.  As long
as applications use the library on a single machine, this
implementation behaves as expected.  However, if the library is used
in the context of a more sophisticated application, say one whose
computation is distributed among a collection of machines, its
behavior can become significantly harder to understand.  In
particular, a distributed implementation might wish to
\emph{replicate} the counter state on each node to improve response
time or fault tolerance.  Unfortunately, adding replication doesn't
come for free.  Attempting to update every replicated copy atomically
is problematic in the absence of sophisticated transaction support,
which impose significant performance penalties.  But, without such
heavyweight mechanisms, applying an \C{Add} operation on one node may
not be instantaneously witnessed on another, which may be in the
process of simultaneously attempting to perform its own \C{Add} or
\C{Mult} action.  While synchronizing the activities of all nodes to
ensure at most one such operation is performed at a time is
impractical, designating a single node to hold the counter state
eschewing replication altogether (as in a typical client-server
configuration), is also an undesirable solution, given the sensitivity
of such architectures to network partitions and server failures, and
the negative performance impact it incurs in geo-distributed
environments~\cite{Walter}.  Removing coordination altogether by
simply replicating counter state without having any supporting
consistency protocol is an equally infeasible approach.

@@

	\vspace*{-1in}

@@

Our main innovation is thus the development of a programming model
that realizes a monadic version control system centered around data,
rather than file, coherence.  The \name (Versioned ML) programming
model comprises the {\sc vst} monad which lets programmers write and
compose concurrent computations around multiple (implicit) versions of
a mergeable datatype, an ordinary ML datatype additionally equipped
with a \emph{merge} function responsible for deriving a consistent
global state from a collection of local versions of that state.  A
computation progresses by forking (i.e., replicating) new versions of
existing versions, allowing synchronization-free local computation to
proceed on those nodes, creating new versions along an existing
\emph{branch} that represents updated local instances, and by merging
branches to realize global consistency.

Our model allows ML programmers to get the benefits of achieving
highly available (low-latency) distributed computation, while
continuing to enjoy the comforts of high-level reasoning and the
familiarity of using standard libraries already provided by the
language.  Notably, while \name\ makes no explicit reference to any
specific operational manifestation of a distributed system (e.g.,
programmers do not need to explicitly manage replicas), we demonstrate
that it can be nonetheless efficiently realized on existing real-world
geo-replicated distributed systems.

@@

    \item We describe \name, a programming model that extends ML to
      brings to bear the power and flexibility of a version control
      system to the administration of replicated data. \name hides the
      complexity of version control behind a monad, and exposes its
      functionality via a simple API that lets programmers define and
      compose distributed ML computations around such data.  Issues
      related to replication only manifest implicitly in the
      definition of a merge operation that defines coherence among
      different instances of (presumably) replicated state.

### 2017-02-28 paper/meta.tex
edits

The problem of multiple LCAs also arises in the context of source
control systems, which employ \emph{ad hoc} mechanisms to pave the way
for three-way merging.  Git~\cite{git}, for instance,
recursively merges LCAs by default to compute a virtual ancestor, which then
serves as the LCA for merging concurrent versions. This method is
demonstrated in the branching structure for a mergeable, replicated
counter as shown in Fig.~\ref{fig:criss-cross-lcas}, where LCAs 5 and
4 of 0 and 3 are merged (with their LCA being 10) to generate -1 as
the virtual LCA to merge 5 and 4. A major downside with this approach
is that it makes no guarantees on the relationship between the virtual
ancestor and its concurrent versions; the former may not even be a
legal ancestor of the latter as per the semantics of the data type.
For instance, suppose the integer type in
Fig.~\ref{fig:criss-cross-lcas} represents a bank account balance,
which is expected to disallows any activity on the account if the
balance is ever known to be less than zero.  From the perspective of
the library designer and its clients, there is no meaningful scenario
in which versions 3 and 0 can emerge from -1, since the only
transition allowed by the semantics from -1 is to itself.  Clearly,
\emph{ad hoc} mechanisms like this are error-prone and difficult to
apply in general.

Fortunately, unlike source control systems where branching structure
is entirely dictated by the user, \name abstracts away branching
structure from the programmer, and hence retains the ability to
manifest it in a way that it deems fit. In particular, \name solves
the problem of multiple LCAs by suitably constraining the branching
structure such that the problem never arises. The constraints are
imposed either implicitly, as a result of how operational semantics
defines an atomic step, or explicitly, by insisting that certain
conditions be met before merging a pair of versions
(\rulelabel{E-Pull-Wait}). Firstly, the operational semantics already
disables criss-cross merges since it only ever merges versions that
are latest on their respective branches. Secondly, we impose certain
pre-conditions on the merging branches to preempt the structure shown
in Fig.~\ref{fig:external-lcas}. The intuition is as follows: consider
the branch $b_3$ at the instance of merging $v_{12}$. Since it has
already merged $v_{22}$, $v_{22}$ could be a common ancestor for $b_3$
and some other branch (call it $b$). Now, if $b_3$ merges $v_{12}$,
same could be true of $v_{12}$. Since $v_{12}$ and $v_{22}$ are not
orderd by the ancestor relation, both become LCAs of
$b_3$ and $b$. We observe that this scenario can be prevented if, when
merging $b_1$, $b_3$ insists on an ancestor relation between the last
merged version ($v_{22}$) and the currently merging version.  We call
the last merged version of a branch its \emph{locus}. By maintaining a
total order of among locii of a branch $b$ as new versions are
created, we effectively guarantee the presence of a version (the
locus) that is lower than all the ancestors of the latest version on
$b$. Next, by requiring the locus to be an ancestor of the merging
version, we ensure that all the common ancestors have a single lowest
element, thus enforcing the uniqueness of the LCA.

@@

\begin{definition} [\bfseries Locus]
Given a branch $b$ and a version $v\in b$, locus ($v_o$) of $v$ is an
ancestor that is not an ancestor of any other external ancestor of
$v$. That is, $\under{H}{v_o \preceq v}$, and there does not exist a
$v_o' \not\in b$ such that $\under{H}{v_o' \preceq v}$ and
$\under{H}{v_o \preceq v_o'}$. We lift the notion of locus to the
level of branches by defining the locus of a branch as the locus of
its latest version.
\end{definition}

In a legal branching history, every version has a unique locus
(property follows from Def.~\ref{def:mergeability} below). Let
\emph{causal history} of a version be defined as the set of all
ancestors of that version. A locus ($v_o$) of a version ($v$) is
therefore the maximum element (lowest element in the ancestor
relation) of the causal history of $v$'s internal ancestor that was
the result of the last merge. We make use of this observation while
defining mergeability.

@@

\begin{theorem} [\bfseries Progress]
In a legal branching history $H$ produced by the operational
semantics, if two branches, $b_i$ and $b_j$ are not mergeable (as per
Def.~\ref{def:mergeability}), then there exist a sequence of fork and
merge operations (between mergeable versions) that can be performed on
$H$ to yeild a new history $H'$, where $b_i$ and $b_j$ are mergeable.
\end{theorem}

Proof is by the induction on the size of branching history. The
observation that the smallest history that is the union of causal
histories of two unmergeable versions (call them $v_i$ and $v_j$) is
smaller than the history that includes the unmergeable versions. Since
causal histories have maximum versions (respective locii), which, by
inductive hypothesis, can be made mergeable, we merge them to yeild a
new version $v$ that is the maximum of the causal histories of both
$v_i$ and $v_j$. If $v_i$ and $v_j$ are the latest versions on their
respective branches, the branches are now mergeable.

>>>>

The problem of multiple LCAs also arises in the context of source
control systems, which employ \emph{ad hoc} mechanisms to pave the way
for three-way merging.  Git~\cite{git}, for instance,
recursively merges LCAs by default to compute a virtual ancestor, which then
serves as the LCA for merging concurrent versions. This method is
demonstrated in the branching structure for a mergeable, replicated
counter as shown in Fig.~\ref{fig:criss-cross-lcas}, where LCAs 5 and
4 of 0 and 3 are merged (with their LCA being 10) to generate -1 as
the virtual LCA to merge 5 and 4. A major downside with this approach
is that it makes no guarantees on the relationship between the virtual
ancestor and its concurrent versions; the former may not even be a
legal ancestor of the latter as per the semantics of the data type.
For instance, suppose the integer type in
Fig.~\ref{fig:criss-cross-lcas} represents a bank account balance,
which is expected to disallow any activity on the account if the
balance is ever known to be less than zero.  From the perspective of
the library designer and its clients, there is no meaningful scenario
in which versions 3 and 0 can emerge from -1, since the only
transition allowed by the semantics from -1 is to itself.  Clearly,
\emph{ad hoc} mechanisms like this are error-prone and difficult to
apply in general.

Fortunately, unlike source control systems where branching structure
is entirely dictated by the user, \name abstracts away branching
structure from the programmer, and hence retains the ability to
manifest it in a way that it deems fit. In particular, \name solves
the problem of multiple LCAs by suitably constraining the branching
structure such that the problem never arises. The constraints are
imposed either implicitly, as a result of how the operational semantics
defines an atomic step, or explicitly, by insisting that certain
conditions be met before merging a pair of versions
(\rulelabel{E-Pull-Wait}). First, the operational semantics already
disables criss-cross merges since it only ever merges versions that
are latest on their respective branches. Second, we impose certain
pre-conditions on the merging branches to preempt the structure shown
in Fig.~\ref{fig:external-lcas}. The intuition is as follows: consider
the branch $b_3$ at the instance of merging $v_{12}$. Since it has
already merged $v_{22}$, $v_{22}$ could be a common ancestor for $b_3$
and some other branch (call it $b$). Now, if $b_3$ merges $v_{12}$,
the same could be true of $v_{12}$. Since $v_{12}$ and $v_{22}$ are not
orderd by the ancestor relation, both become LCAs of
$b_3$ and $b$. We observe that this scenario can be prevented if, when
merging $b_1$, $b_3$ insists on an ancestor relation between the last
merged version ($v_{22}$) and the currently merging version.  We call
the last merged version of a branch its \emph{locus}. By maintaining a
total order of among locii of a branch $b$ as new versions are
created, we effectively guarantee the presence of a version (the
locus) that is lower than all the ancestors of the latest version on
$b$. Next, by requiring the locus to be an ancestor of the merging
version, we ensure that all the common ancestors have a single lowest
element, thus enforcing the uniqueness of the LCA.

@@

\begin{definition} [\bfseries Locus]
Given a branch $b$ and a version $v\in b$, the locus ($v_o$) of $v$ is an
ancestor that is not an ancestor of any other external ancestor of
$v$. That is, $\under{H}{v_o \preceq v}$, and there does not exist a
$v_o' \not\in b$ such that $\under{H}{v_o' \preceq v}$ and
$\under{H}{v_o \preceq v_o'}$. We lift the notion of locus to the
level of branches by defining the locus of a branch as the locus of
its latest version.
\end{definition}

In a legal branching history, every version has a unique locus
(property follows from Def.~\ref{def:mergeability} below). We
define the \emph{causal history} of a version as the set of all
ancestors of that version. A locus ($v_o$) of a version ($v$) is
therefore the maximum element (lowest element in the ancestor
relation) of the causal history of $v$'s internal ancestor that was
the result of the last merge. We make use of this observation while
defining mergeability.

@@

\begin{theorem} [\bfseries Progress]
In a legal branching history $H$ produced by the operational
semantics, if two branches, $b_i$ and $b_j$ are not mergeable (as per
Def.~\ref{def:mergeability}), then there exists a sequence of fork and
merge operations (between mergeable versions) that can be performed on
$H$ to yeild a new history $H'$, where $b_i$ and $b_j$ are mergeable.
\end{theorem}

The proof is by the induction on the size of branching history. The
key observation is that the smallest history that is the union of
causal histories of two unmergeable versions (call them $v_i$ and
$v_j$) is smaller than the history that includes the unmergeable
versions. Since causal histories have maximum versions (respective
locii), which, by the inductive hypothesis, can be made mergeable, we
merge them to yield a new version $v$ that is the maximum of the
causal histories of both $v_i$ and $v_j$. If $v_i$ and $v_j$ are the
latest versions on their respective branches, the branches are now
mergeable.

### 2017-02-28 paper/model.tex
edits

The \name programming model lets programmers define and compose
concurrent computations around versioned data structures.
Fig.~\ref{fig:dali-monad} shows the signature of the \name module that
implements the programming model along the lines of the well-known
\C{State} monad~\cite{wadler-monad}. The monad encapsulates a
versioned functional state (\C{'a}) and the type \C{('a, 'b) t}
represents a monadic computation that returns a \C{'b} result.
Functions \C{return} and \C{bind} have their usual monadic
\begin{figure}
\begin{center}
  \begin{ocaml}
  module type VST = sig
    type ('a, 'b) t
    val return : 'b -> ('a, 'b) t
    val bind : ('a, 'b) t -> ('b -> ('a, 'c) t) -> ('a, 'c) t
    val get_current_version: unit -> ('a, 'a) t
    val with_init_version_do: 'a -> ('a, 'b) t -> 'b
    val fork_version : ('a, 'b) t -> ('a, unit) t
    val sync_next_version: unit -> 'a -> ('a, 'a) t
  end
  \end{ocaml}
  \caption{Signature of the \name monad}
  \label{fig:dali-monad}
\end{center}
\end{figure}
interpretation. \C{get\_current\_version} is like the \C{State}
monad's \C{get}; it returns the versioned state encapsulated by the
monad. To initiate a computation, we use \C{with\_init\_version\_do}
which runs a monadic computation against a given initial version and
returns the result. The \C{fork\_version} operation returns a
computation that forks a new concurrent version from the current
version, and runs the given monadic computation asynchronously against
the forked version.  \C{sync\_next\_version} (simply called \C{sync})
accepts a \emph{proposal} for the next version of the state; this
proposal is the current local version of the state that reflects local
modifications not yet witnessed by any other concurrently executing
computation.  The operation returns (via a monad) the actual next
version, which becomes the current version for the rest of the
computation.  This version is created by merging the proposal with a
subset of concurrent versions that have become available since the
last merge or fork. Thus, \C{sync} effectively lets a computation
synchronize with a subset of concurrent computations to obtain their
latest updates.

>>>>

\name\ lets programmers define and compose concurrent computations
around versioned data structures.  Fig.~\ref{fig:dali-monad} shows the
signature of the \nameMonad module that implements the programming
model along the lines of the well-known \C{State} monad. The monad
encapsulates a versioned functional state (\C{'a}) and the type
\C{('a, 'b) t} represents a monadic computation that returns a \C{'b}
result.  Functions \C{return} and \C{bind} have their usual monadic
\begin{figure}
\begin{center}
  \begin{ocaml}
  module type VST = sig
    type ('a, 'b) t
    val return : 'b -> ('a, 'b) t
    val bind : ('a, 'b) t -> ('b -> ('a, 'c) t) -> ('a, 'c) t
    val get_current_version: unit -> ('a, 'a) t
    val with_init_version_do: 'a -> ('a, 'b) t -> 'b
    val fork_version : ('a, 'b) t -> ('a, unit) t
    val sync_next_version: unit -> 'a -> ('a, 'a) t
  end
  \end{ocaml}
  \caption{Signature of the \name monad}
  \label{fig:dali-monad}
\end{center}
\end{figure}
interpretation. \C{get\_current\_version} is like the \C{State}
monad's \C{get}; it returns the versioned state encapsulated by the
monad. To initiate a computation, we use \C{with\_init\_version\_do}
which runs a monadic computation against a given initial version and
returns the result. The \C{fork\_version} operation returns a
computation that forks a new concurrent version from the current
version, and runs the given monadic computation asynchronously against
the forked version.  \C{sync\_next\_version} (simply called \C{sync})
accepts a \emph{proposal} for the next version of the state; this
proposal is the current local version of the state that reflects local
modifications not yet witnessed by any other concurrently executing
computation.  The operation returns (via a monad) the actual next
version, which becomes the current version for the rest of the
computation.  This version is created by merging the proposal with a
subset of concurrent versions that have become available since the
last merge or fork. Thus, \C{sync} effectively lets a computation
synchronize with a subset of concurrent computations to obtain their
latest updates.
